Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-09-04 papers/paper/ms.tex
update ms

Despite this and other observational challenges, young close-in planet
discovery has matured over the past decade, primarily due to Kepler,
K2, TESS, and Gaia
\citep[e.g.][]{Meibom_2013,Mann_K2_25_2016,Curtis_2018,Livingston_2018,Bouma_2020_toi837,Plavchan_2020,Newton_2021,Nardiello_2022,Barber_2022,Zhou_2022,Zakhozhay_2022,Wood_2023}.
The strategy pursued by most groups during the 2010s was to focus on
stars with known ages---obvious members of open clusters---and to
search them for transiting planets.  The resulting stellar, and
assumed planetary, ages are precise at the $\approx$10\% level.  
An alternative strategy, facilitated by Gaia, is to select \deleted{transiting}
planets based on empirical youth indicators, and to then search their local
neighborhoods for comoving \added{stellar }companions \citep[e.g.][]{Tofflemire_2021}.

While stellar ensembles provide the gold standard for astronomical ages,
this ``cluster-first'' approach has a major limitation: most 
stars are in the field.  Only $\lesssim$1\% of stars within
$\approx$500\,pc have been associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
The implication for known planetary ages can appreciated by querying the NASA Exoplanet Archive
\citep[NEA;][]{2013PASP..125..989A} for transiting planets younger
than 1\,Gyr.  Requiring $t$$<$1\,Gyr at 2$\sigma$ precision gives
$\approx$50 such planets at the time of writing.  Most of these young
planets are in clusters, and were found by Kepler, K2, or TESS.
However, these surveys have cumulatively discovered \deleted{a total of }$\approx$5{,}000 planets.
Assuming a constant star formation history, we would expect an
order of magnitude more sub-gigayear transiting planets\added{ than the 50 currently known}.

@@

\begin{enumerate}[label={\it \roman*)},leftmargin=12pt,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item The Kepler Object of Interest catalog, and our vetting of
    false positives within it, has now reached maturity
    \citep[e.g.][]{Thompson_2018}.
  %
  \item Measured rotation periods of FGK stars in open clusters now
    show not only the average evolution of $P_{\rm rot}(T_{\rm
    eff},t)$, but also how the astrophysical dispersion of stars
    around this average converges by the $\approx$700\,Myr age of
    Praesepe
    \citep[e.g.][]{Curtis_2019_ngc6811,Gillen_2020,Rampalli_2021,Fritzewski_2021,Rebull_2022,Dungee_2022,2023AJ....166...14B}.
  %
  \item Using open cluster data, we can marginalize over the range of
    ages that might explain any one star's rotation period
    \citep{Bouma_2023}.  This represents an improvement in the
    accuracy of uncertainty propagation relative to previous
    calibrations.
  %
  \item We can identify \replaced{binary stars}{stars that are
    binaries and that have high metallicity} with greater fidelity\deleted{ than
    in the past}, which can clarify otherwise problematic estimates of
    rotation-based ages.
  %
  \item New open clusters in various stages of dissolution have been
    found in the Kepler field
    \cite[e.g.][]{2019AJ....158..122K,Kerr_2021,Barber_2022}.
    These clusters offer stars that we can use to test
    the reliability of \replaced{the available single-star}{our} age-dating methods.
\end{enumerate}

>>>>

Despite this and other observational challenges, young close-in planet
discovery has matured over the past decade, primarily due to Kepler,
K2, TESS, and Gaia
\citep[e.g.][]{Meibom_2013,Mann_K2_25_2016,Curtis_2018,Livingston_2018,Bouma_2020_toi837,Plavchan_2020,Newton_2021,Nardiello_2022,Barber_2022,Zhou_2022,Zakhozhay_2022,Wood_2023}.
The strategy pursued by most groups during the 2010s was to focus on
stars with known ages---obvious members of open clusters---and to
search them for transiting planets.  The resulting stellar, and
assumed planetary, ages are precise at the $\approx$10\% level.  
An alternative strategy, facilitated by Gaia, is to select \deleted{transiting }planets
based on empirical youth indicators, and to then search their local
neighborhoods for comoving \added{stellar }companions \citep[e.g.][]{Tofflemire_2021}.

While stellar ensembles provide the gold standard for astronomical ages,
this ``cluster-first'' approach has a major limitation: most 
stars are in the field.  Only $\lesssim$1\% of stars within
$\approx$500\,pc have been associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
The implication for known planetary ages can be appreciated by querying the NASA Exoplanet Archive
\citep[NEA;][]{2013PASP..125..989A} for transiting planets younger
than 1\,Gyr.  Requiring $t$$<$1\,Gyr at 2$\sigma$ precision gives
$\approx$50 \deleted{such }planets at the time of writing.  Most of these young
planets are in clusters, and were found by Kepler, K2, or TESS.
However, these surveys have cumulatively discovered \deleted{a total of }$\approx$5{,}000 planets.
Assuming a constant star formation history, we would expect an
order of magnitude more sub-gigayear transiting planets\added{ than the 50 currently known}.

@@

\begin{enumerate}[label={\it \roman*)},leftmargin=12pt,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item The Kepler Object of Interest catalog, and our vetting of
    false positives within it, has now reached maturity
    \citep[e.g.][]{Thompson_2018}.
  %
  \item Measured rotation periods of FGK stars in open clusters now
    show not only the average evolution of $P_{\rm rot}(T_{\rm
    eff},t)$, but also how the astrophysical dispersion of stars
    around this average converges by the $\approx$700\,Myr age of
    Praesepe
    \citep[e.g.][]{Curtis_2019_ngc6811,Gillen_2020,Rampalli_2021,Fritzewski_2021,Rebull_2022,Dungee_2022,2023AJ....166...14B}.
  %
  \item Using open cluster data, we can marginalize over the range of
    ages that might explain any one star's rotation period
    \citep{Bouma_2023}.  This represents an improvement in the
    accuracy of uncertainty propagation relative to previous
    calibrations.
  %
  \item We can identify \replaced{binary stars}{stars that are
    binaries and that have high metallicity} with greater fidelity\deleted{ than
    in the past}, which can clarify otherwise problematic estimates of
    rotation-based ages.
  %
  \item New open clusters in various stages of dissolution have been
    found in the Kepler field
    \cite[e.g.][]{2019AJ....158..122K,Barber_2022,Kerr2024}.
    These clusters offer stars that we can use to test
    the reliability of \replaced{the available single-star}{our} age-dating methods.
\end{enumerate}
